Model Finding for Recursive Functions in SMT\thanks{ This research is partially supported by the Inria technological development action ``Contre-exemples utilisables par Isabelle et Coq'' (CUIC). }

### 2016-01-25 ijcar2016/body.tex
tuned terminology

\begin{definition}\rm
A formula $\varphi$ is in \emph{definitional form with respect to}
$\{ \con{f}_1, \ldots, \con{f}_n \} \subseteq \sfuns{\Sigma}$ if it is of the
form
%
%\kern-\abovedisplayskip
%\kern+\abovedisplayshortskip
%
$(\forallf{\con{f}_1} \vec x_1.\; \varphi_1) \wedge \cdots \wedge
(\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \psi$,
%
where $\con{f}_1, \ldots, \con{f}_n$ are distinct function symbols,
%ct added
$\forallf{\con{f}_i} \vec x_i.\; \varphi_i$ is a function definition
for $i = 1, \ldots, n$,
%
%\ct{I think we need more conditions here on how a symbol $\con{f}_i$ can occur in $\varphi_j$.}
%\jb{do we? I thought that the admissibility criterion below is enough.}
% should be enough, unless I'm missing something? -ajr
and $\psi$ contains no function definitions.
We call $\psi$ the \emph{main formula} of $\varphi$.
%\ct{Don't we need additional restrictions on the $\varphi_i$'s here?
%(E.g., they are quantifier-free, or perhaps existential or containing only let binders.)}
% Now fixed above -ajr
\end{definition}

@@

The intuition of the above proof is as follows.
First, $\conv( \varphi )$ cannot be unsatisfiable when $\varphi$ is satisfiable
since any $\Sigma$-interpretation that satisfies $\varphi$ can be extended in a straightforward way to 
an $\extendsig{\Sigma}$-interpretation that satisfies $\conv( \varphi )$, by
interpreting the abstract types in the same way as the cartesian products they
abstract, thereby satisfying all existential constraints introduced by $\conv$.
Second, if a model is found for $\conv( \varphi )$,
then the constraints $\exists a : \fargtype{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t$
occurring in $\conv( \varphi )$ ensure that this model also satisfies
a $\Sigma$-formula that is closed under function expansion and
that entails the main formula of $\varphi$.
This implies the existence of a model for $\varphi$,
assuming $\DDD$ is admissible.
%For clarity, we demonstrate the consequences of this theorem by means of an example.

>>>>

\begin{definition}\rm
A formula $\varphi$ is in \emph{definitional form with respect to}
$\{ \con{f}_1, \ldots, \con{f}_n \} \subseteq \sfuns{\Sigma}$ if it is of the
form
%
%\kern-\abovedisplayskip
%\kern+\abovedisplayshortskip
%
$(\forallf{\con{f}_1} \vec x_1.\; \varphi_1) \wedge \cdots \wedge
(\forallf{\con{f}_n} \vec x_n.\; \varphi_n) \wedge \psi$,
%
where $\con{f}_1, \ldots, \con{f}_n$ are distinct function symbols,
%ct added
$\forallf{\con{f}_i} \vec x_i.\; \varphi_i$ is a function definition
for $i = 1, \ldots, n$,
%
%\ct{I think we need more conditions here on how a symbol $\con{f}_i$ can occur in $\varphi_j$.}
%\jb{do we? I thought that the admissibility criterion below is enough.}
% should be enough, unless I'm missing something? -ajr
and $\psi$ contains no function definitions.
We call $\psi$ the \emph{goal} of $\varphi$.
%\ct{Don't we need additional restrictions on the $\varphi_i$'s here?
%(E.g., they are quantifier-free, or perhaps existential or containing only let binders.)}
% Now fixed above -ajr
\end{definition}

@@

The intuition of the above proof is as follows.
First, $\conv( \varphi )$ cannot be unsatisfiable when $\varphi$ is satisfiable
since any $\Sigma$-interpretation that satisfies $\varphi$ can be extended in a straightforward way to 
an $\extendsig{\Sigma}$-interpretation that satisfies $\conv( \varphi )$, by
interpreting the abstract types in the same way as the cartesian products they
abstract, thereby satisfying all existential constraints introduced by $\conv$.
Second, if a model is found for $\conv( \varphi )$,
then the constraints $\exists a : \fargtype{\con{f}}.\;\, \vecfarg{\con{f}}( a ) \teq \vec t$
occurring in $\conv( \varphi )$ ensure that this model also satisfies
a $\Sigma$-formula that is closed under function expansion and
that entails the goal of $\varphi$.
This implies the existence of a model for $\varphi$,
assuming $\DDD$ is admissible.
%For clarity, we demonstrate the consequences of this theorem by means of an example.
